---
output: 
  github_document:
    toc: true
    html_preview: false
params: 
    channel: bus
title: "`r str_to_title(params$channel)`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# import libraries up here so we don't print out any warnings that make our document messy
library(tidyverse)
library(corrplot)
library(caret)
library(gbm)
library(leaps)
library(caret)
library(cowplot)
```

## Intro
This project aims at building predictive models on the [Online News Popularity](https://archive.ics.uci.edu/ml/datasets/Online+News+Popularity#) dataset. Our goal is to build models that predict the number of shares an article receives based on characteristics of the article. This work is important to help writers/companies understand factors that influence article success (as measured by the number of shares an article obtains). This work could also be integrated in the calculation of advertisement space on articles (articles that receive more shares should demand more for ad space). 

The characteristics we will explore include day of week, number of links, word count of title, word count of content

+ `weekday`: day of week that article was published (Monday, Tuesday, ...)
+ `num_hrefs`: number of links referenced in article
+ `n_tokens_title`: word count of title
+ `n_tokens_content`: word count of article
+ `rate_positive_words` and `rate_negative_words`: rate of positive/negative words among non-neutral tokens

The models used to build ... continue from here
Data was split by channel type...

## Requirements
```{r 0_requirements, eval=FALSE}
library(tidyverse)
library(corrplot)
```

## Data manipulaton

### Read in Data
```{r 1_readData}
# read entire dataset
  # subset to get the data channel we want
    # remove unnessary columns
df <- read_csv('data/OnlineNewsPopularity.csv') %>%
        filter((!!sym(paste0('data_channel_is_', params$channel))) == 1) %>%
          select(-contains('data_channel_is'))

dim(df)
```

The object `df` now holds a subset of the data according to the specified data channel. Now, we must create new features to consolidate variables and prepare the dataset for exploratory data analysis:

```{r}
# Create a variable that holds day of week
levels = c('Sunday', 'Monday', 'Tuesday', 'Wednesday','Thursday', 'Friday', 'Saturday')
df <- df %>% mutate(weekday = ifelse(weekday_is_monday==1, 'Monday', 
                                ifelse(weekday_is_tuesday==1, 'Tuesday',
                                ifelse(weekday_is_wednesday==1, 'Wednesday',
                                ifelse(weekday_is_thursday==1, 'Thursday', 
                                ifelse(weekday_is_friday==1, 'Friday',
                                ifelse(weekday_is_saturday==1, 'Saturday', 'Sunday'
                                       ))))))) %>%
                mutate(weekday = factor(weekday, levels = levels))


# function to create the popularity column. 
# popularity rating was created with summary stat info (25%, median, and 75%)
popularityCol <- function(dataset){
  dataset <- dataset %>% 
    mutate("Popularity" = if_else(shares > 2800, "Very popular",
                           if_else(shares > 1400, "Somewhat popular", 
                            if_else(shares > 946, "Not too popular", "Not at all popular")) 
                           )
    )            
  return(dataset)
}

# Data set using popularityCol function.
df <- popularityCol(df) %>% as_tibble()

# Overwrite popularity column with factor version
df$Popularity <- as.factor(df$Popularity)

# Use ordered function on a factor to order the levels
df$Popularity <- ordered(df$Popularity, levels = c("Not at all popular", "Not too popular", "Somewhat popular", "Very popular"))


# Deleting variables
#df <- df[-1:2, 34:38]
```

Our data set contains 30 000 instances; each consists of 60 attributes and one dependent variable, the number of shares of the article.

The following variables can be omitted:

url: URL of the article (non-predictive)
timedelta: Days between the article publication and the dataset acquisition (non-predictive)
five LDA variables
is_weekend, since it seems to be duplicating days of week
kw_min_min, kw_avg_min, kw_min_avg have a number of negative values


We now split the data into train and test sets for predicitve modeling.

```{r}
set.seed(123)
# split data into train and test sets
train_rows <- sample(nrow(df), nrow(df)*0.7)
trainData <- df[train_rows,]
testData <- df[-train_rows,] 
```

## Summarizations

We will use train set for EDA. 

### Exploratory Data Anaysis

The summary statistics of targeted variable (shares)

```{r}
# summary statistics
share_stat <- trainData %>% 
                summarise(Count = n(),
                          Min = min(shares), 
                          Q1 = quantile(shares, 0.25),
                          Median = median(shares),
                          Average = mean(shares),
                          Q3 = quantile(shares, 0.75),
                          Max = max(shares),
                          Std.Dev = sd(shares)
                          )

# Display a table of the summary stats.
knitr::kable(share_stat, caption = "Summary Stats by shares", digits = 2)

```

### Shares by days of week

```{r}
trainData %>%
  group_by(weekday) %>%
  summarise(total_shares = sum(shares), avg_shares = round(mean(shares)), max_shares = max(shares)) %>%
  knitr::kable()

```

The above table shows a breakdown of total, average, and maximum number of shares for articles published on a specific weekday for this channel. Some channels tend to have more popular days than others.

### shares by popularity

```{r}
trainData %>% 
  group_by(Popularity) %>%
  summarise(total_shares = sum(shares), avg_shares = round(mean(shares)), max_shares = max(shares)) %>%
  knitr::kable()
```

The above table show a summary of the newly created `popularity` variable.

### count of news by popularity over different days of week

```{r}
#Bar plot of weekday by popularity 

ggplot(data = trainData, aes(x = weekday)) +
  geom_bar(aes(fill = as.factor(Popularity))) + 
  labs(x = "Days of week", 
       title = "Days of week by popularity") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  scale_fill_discrete(name = "Popularity") 
            
```

The bar plot above shows a breakdown of the number of articles published vs day of week. We can also see a breakdown of the proportion of popularity of articles for each day. This plot is important to understand if a greater number of shares exhibited by a day of the week is due to more articles being published on that day, or if it is due to that day of week having a direct effect on the number of shares.

For example, if all days have the same number of 'very popular' articles, we could hypothesize that day of week does not have an effect on producing 'very popular' articles.


### Days of week

```{r 2_eda}
# histogram for day of week vs shares

g1 <- trainData %>% ggplot(aes(x=weekday, y=shares)) +
        geom_bar(stat="identity", fill = "darkblue") + 
   theme(axis.text.x = element_text(angle = 45, vjust = .75)) +
        ggtitle('Day of Week and Total Number of Shares')
g1
```

The above histogram shows a basic count of the total number of shares for all articles published on each day of week.

### Number of links

```{r 1_eda}
# simple scatter plot
g2 <- trainData %>% ggplot(aes(x=num_hrefs, y=shares)) +
        geom_point(size=2, shape=23) +
        ylim(0, 10000) +
        ggtitle("Number of links")  
g2
```

In the above scatter, we compare the number of links in an article to its shares. This plot is motivated by the implementation of Google's [PageRank Algorithm](https://en.wikipedia.org/wiki/PageRank).

### Number of link by weekday.
```{r}

ggplot(trainData, aes(x = weekday, 
                                y = num_hrefs, 
                                fill = weekday)
                            ) +
  geom_boxplot() +
  scale_x_discrete("Time weekday") +
  ggtitle("Numbers link comparison by days of week") +
  scale_fill_brewer(palette = "BuPu") +
  scale_fill_discrete(name = "Weekday") 

```


### Number of words in the title and content

```{r}
library(cowplot)

#scatter plots of Number of words in the title
g3 <- ggplot(data = trainData, aes(x =  n_tokens_title, 
                      y = shares)) +
      geom_point(alpha = 0.50) + 
  #theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
      ggtitle("Word count in the title")  


#scatter plots of Number of words in the content
g4 <- ggplot(data = trainData, aes(x =  n_tokens_content, 
                      y = shares)) +
      geom_point(alpha = 0.50) + 
  #theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
      ggtitle("Word count in the content")

plot_grid(g3, g4,  labels = c('A', 'B'))   
```
Graph A shows the number of words in the title compared to the number of shares. Perhaps a quadratic relationship is appropriate for this variable if a bell shape appears in the plot.
Graph B shows a comparison of the number of words that appear in the article compared to the number of shares. Perhaps a negative linear relationship is appropriate if the data exihibits a slightly negative slops.

### Unique words count

Below, we see a plot of the number of unique words in each article compared to the number of shares. Across channels, there does not seem to be a strong linear correlation between these two variables. However, this variable may have significance in tree based models.

```{r}
ggplot(data = trainData, aes(x =  n_unique_tokens, 
                      y = shares)) +
      geom_point(alpha = 0.50) + 
  #theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
      ggtitle("Unique Word count") +
      ylim(0, 10000)

```

### Number of image and video

Below, we compare the number of images/videos to the number of shares. Across channels, it seems that most articles have less than 5 images, which makes fitting a linear model difficult to this variable.

```{r}
#scatter plots of Number of words in the content
g5 <- ggplot(data = trainData, aes(x =  num_imgs, 
                      y = shares)) +
      geom_point(alpha = 0.50) +
  #theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
      ggtitle("Number of image") +
      ylim(0, 10000)


#scatter plots of Number of words in the content
g6 <- ggplot(data = trainData, aes(x =  num_videos, 
                      y = shares)) +
      geom_point(alpha = 0.50) + 
  #theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
      ggtitle("Number of video") +
      ylim(0, 10000)


plot_grid(g5, g6,  labels = c('C', 'D')) 
```
### Number of keywords

Below, we compare the number of keywords that appear in an article to the number of shares. Across channels, there does not appear to be much linear correlation between the two variables.

```{r}
ggplot(data = trainData, aes(x =  num_keywords, 
                      y = shares)) +
      geom_point(alpha = 0.50) + 
  #theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
      ggtitle("Number of keywords")

```

### Number of positive and negative words rate

Below we compare the positive/negative rate of words that appear in an article to the number of shares of an article. Across channels, there seems to be a slight positive correlation between positive words rate and shares. Additionally, there seems to be a slight negative correlation between negative words rate and shares.

```{r}
#scatter plots of positive and negative words rate
g7 <- ggplot(data = trainData, aes(x =  rate_positive_words, 
                      y = shares)) +
      geom_point(alpha = 0.50) +
  #theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
      ggtitle("Positive words rate")


#scatter plots of Number of words in the content
g8 <- ggplot(data = trainData, aes(x =  rate_negative_words, 
                      y = shares)) +
      geom_point(alpha = 0.50) +
  #theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
      ggtitle("Negative words rate")


plot_grid(g7, g8,  labels = c('A', 'B')) 
```

### Title subjectivity

Below we compare title subjectivity to the number of shares. Across channels, there does not seem to be a storng linear correlation between the two variables.

```{r}

#scatter plot of title subjectivity
ggplot(data = trainData, aes(x =     title_subjectivity, 
                      y = shares)) +
      geom_point(alpha = 0.50) +
  #theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
      ggtitle("title subjectivity")

```

### Number of positive and negative words rate

```{r}
#scatter plots of positive and negative words rate
g9 <- ggplot(data = trainData, aes(x =  avg_positive_polarity, 
                      y = shares)) +
      geom_point(alpha = 0.50) +
  #theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
      ggtitle("Average positive polarity")


#scatter plots of Number of words in the content
g10 <- ggplot(data = trainData, aes(x =  avg_negative_polarity, 
                      y = shares)) +
      geom_point(alpha = 0.50) +
  #theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
      ggtitle("Average negative polarity")


plot_grid(g9, g10,  labels = c('A', 'B')) 
```

### Correlation with numeric variables

David- We can add more numerical variables here. what do you think?
```{r}
df_tmp <- trainData %>% select(c('n_tokens_title', 
                          'n_tokens_content',
                          'n_unique_tokens',
                          'num_hrefs',
                          'num_imgs',
                          'num_videos',
                          'num_keywords',
                          'is_weekend',
                          'rate_positive_words',
                          'rate_negative_words',
                          'title_subjectivity',
                          'avg_positive_polarity',
                          'avg_negative_polarity',
                          'shares',))
corrplot(cor(df_tmp), type = 'lower', diag = FALSE)

```

Above shows the correlation matrix for other numerical variables. Shares is the bottom row. We use this plot to find other variables that might have weak correlation with shares and make sure to include these in our model building phase.

## Modeling

### Linear Regression

Linear regression is one of the commonly used supervised methods for modeling and useful tool for predicting a quantitative response on the basis of a single or multiple predictor variables. The idea of linear regression is that the model finds the best fit line between predictor variables and response variable, minimizing sum of squared errors. Also, Linear Regression is flexible in which to include many explanatory variables, higher order terms, and/or interaction, so we can see the effect the combinations on prediction.  Here we will fit two different linear regression models.   

#### Linear model 1
The first linear regression model will have predictors selected by stepwise selection. After choosing the subset of predictors, we will use repeated cross-validation with 10 folder and will find the RMSE and R2.

```{r}
# Stepwise model selection
lmFitSelect <- lm(shares ~ n_tokens_title + n_tokens_content+ is_weekend +  num_hrefs+  num_imgs+ num_videos+num_keywords+ rate_positive_words + title_subjectivity  + I(n_tokens_content^2) + I(num_imgs^2)+ I(num_videos^2) + I(num_hrefs^2) + weekday, data = trainData)
models <- step(lmFitSelect, trace=0)
summary(models)
#lmFitSelectFull <- lm(shares ~ ., data = trainData[-57])
#models2 <- step(lmFitSelectFull, trace=0)

```

David - this is the repeat of the first one, but including all the variables. Which one should we use? Do we need to stick with the first one?

```{r}
lmFitSelectFull <- lm(shares ~ ., data = trainData[-c(1:2,57)])
models2 <- step(lmFitSelectFull, trace=0)
summary(models2)
models2
```


```{r}
# train data with variables chosen by stepWise
set.seed(10)
lm.fit1 <- train(shares ~ num_imgs + num_videos + rate_positive_words + I(num_videos^2) + 
    I(num_hrefs^2), data = train_df,
                        method="lm",
                        preProcess = c("center","scale"),
                        trControl = trainControl(method = "repeatedcv", number = 10, repeats = 3))

lm.fit1

```

David - This is again the follow up model fit from all variables. 

```{r}


lm.fit3 <- train(shares ~ n_non_stop_unique_tokens + num_hrefs + 
    num_imgs + num_videos + kw_min_max + kw_max_max + kw_max_avg + 
    kw_avg_avg + LDA_03 + min_negative_polarity, data = trainData[-c(1:2, 
    57)],
                        method="lm",
                        preProcess = c("center","scale"),
                        trControl = trainControl(method = "repeatedcv", number = 10, repeats = 3))

lm.fit3


```

#### Linear model 2 - Logarithmic Linear Regression

Now, let's look at regresssing on the log-transformed target variable of shares.

```{r}
# get all numeric columns
train_df <- trainData[ ,unlist(lapply(trainData, is.numeric))]
test_df <- testData[ ,unlist(lapply(testData, is.numeric))]

# code used to get regression variables
## fit using forward selection
#forward <- regsubsets(log(shares) ~ .,
#                      data = train_df,
#                      nvmax = 10,
#                      method = "forward")
## summary
#mod_summary <- summary(forward)

# train model
lm.fit2 <- lm(log(shares) ~ n_tokens_content + num_hrefs + average_token_length + 
                        num_keywords + kw_min_min + kw_max_avg + kw_avg_avg + 
                        is_weekend + LDA_04 + global_subjectivity, 
          data = train_df
          )

          
lm.fit2

# David- With train function. - It is upto you which one you use. Dr. P was emphasizing using cross validation, so I just added here so that you can choose.

lm.fit4 <- train(log(shares) ~ n_tokens_content + num_hrefs + average_token_length + 
                        num_keywords + kw_min_min + kw_max_avg + kw_avg_avg + 
                        is_weekend + LDA_04 + global_subjectivity, 
          data = train_df,
                        method="lm",
                        preProcess = c("center","scale"),
                        trControl = trainControl(method = "repeatedcv", number = 10, repeats = 3))

lm.fit4

#predict on test data
#predLm2 <- predict(lm.fit2, test_df)

# calculate rmse
#rmseLm2 <- sqrt(mean((predLm2 - test_df$shares)^2))
#rmseLm2
```

### Ensemble Tree-based model

Tree-based method splits up response variable into subsets based on their relationship to one or more predictor variables. Because it is easy to understand and interpret output and no statistical assumptions is necessary, regression tree (continuous variable) and classification tree (group membership) are commonly used. However, these models are influenced vastly even there is a small changes in the data. Therefore, Ensemble tree-based model are highly used in machine learning to gain strong prediction even though it lose interpretability because it get rid of variation of data set to data set and average of predication for a final prediction. In this project, we will use the two of most commonly used ensemble methods: Random Forests and Boosting. 

#### Random Forest Model

Random Forest method,  which shares the idea of bagging, but extends the idea and only include random subset of predictors for each bootstrap sample/tree fit instead of including every predictor in each of the tree.  In doing so, one or two good predictors won’t dominate the tree fit. Random forest method use mtry as tuning parameter and the number of randomly selected predictor is obtained using *m* = *p*/3 formula (p as the number of predictors). Below, you will see the result of training with the random forest method. By choosing randomly selected subset of predictors in each tree, we will possibly reduce the correlation and gain stronger prediction.

```{r}
# get all numeric columns
train_df <- trainData[ ,unlist(lapply(trainData, is.numeric))]
test_df <- testData[ ,unlist(lapply(testData, is.numeric))]
# tuning parameter is mtry, use values of 1,2,..,10
rfFit <- train(shares ~ n_tokens_title + n_tokens_content+
                 n_unique_tokens+avg_positive_polarity+
                 avg_negative_polarity + num_hrefs +  num_imgs +
                 num_videos + num_keywords + title_sentiment_polarity +
                 rate_positive_words + rate_negative_words +
                 title_subjectivity + weekday,
               data = trainData, 
               method = "rf", 
               trControl = trainControl(method = "cv", number = 5),
               preProcess = c("center", "scale"),
               tuneGrid = data.frame(mtry = (1:10)))

rfFit
```

#### Boosted Tree Model

The boosted tree is another tree-based regression model. This model aims to predict the residuals between the number of shares (target variable) of each observation and the average number of shares. To do this, the model builds smaller trees of specified depth that add/subtract to predictions with the hope that predictions move closer to their residual values. To prevent overfitting of data, trees are constricted to a 'shrinkage' parameter (that takes a value between 0 and 1) which limits the amount of boosting on predictions. 

For a really good video explanation, watch [this](https://www.youtube.com/watch?v=3CC4N4z3GJc).

```{r}
# get all numeric columns
train_df <- trainData[ ,unlist(lapply(trainData, is.numeric))]
test_df <- testData[ ,unlist(lapply(testData, is.numeric))]

# declare grid of values to test in cross validation
## code retrieved from https://topepo.github.io/caret/model-training-and-tuning.html
gbmGrid <-  expand.grid(interaction.depth = c(1, 5, 9), # complexity of tree
                        n.trees = (1:30)*50, # number of iterations (i.e. trees)
                        shrinkage = 0.1, # learning rate
                        n.minobsinnode = 20) # minimum number of samples in a node to commence splitting

# train using crossvalidation, print out best fitting parameters
boostFit <- train(shares ~ .,
                data = train_df,
                method = "gbm",
                trControl = trainControl("cv", number = 10),
                verbose = FALSE,
                #tuneGrid = gbmGrid
                )
boostFit$bestTune

# evaluate on test dataset
#boostPred <- predict(boostFit, newdata = dplyr::select(test_df, -shares), n.trees = 100)
#boostRMSE <- sqrt(mean((boostPred-test_df$shares)^2))
#boostRMSE
```

David- are you going to use tuneGrid? You created one, but didn't include in the model. 

## Comparison

Now, we compare the 4 models developed above. Each model was evaluated on a test data set and Root MSE values(roof of test prediction error) were compared. 
```{r}
# Predict on test data

predLm1 <- predict(lm.fit1, newdata = trainData)
predLm2 <- predict(lm.fit2, newdata = test_df)
rfPred <- predict(rfFit, newdata = trainData)
boostPred <- predict(boostFit, newdata = test_df)

# Calculate rmse

rmseLm1 <- sqrt(mean((predLm1 - trainData$shares)^2))
rmseLm2 <- sqrt(mean((predLm2 - test_df$shares)^2))
rfMSE <- sqrt(mean((rfPred - trainData$shares)^2))
boostRMSE <- sqrt(mean((boostPred - test_df$shares)^2))

rmseTotal <- data.frame('Linear Regression Model 1' = rmseLm1, 
                   'Linear Regression Model 2' = rmseLm2, 
                   'Random Forest Model' = rfMSE, 
                   'Boosting Model' = boostRMSE)

knitr::kable(t(rmseTotal),
               digits=3,
               caption="Summary Table of RMSE score",
               col.names = "RMSE")
```

David- I did this part because I will definitely need help on automation if I do it. 