---
output: 
  github_document:
    toc: true
    toc_depth: 4
    html_preview: false
params: 
    channel: bus
title: "`r str_to_title(params$channel)`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)

# import libraries up here so we don't print out any warnings that make our document messy
library(tidyverse)
library(corrplot)
library(caret)
library(gbm)
library(leaps)
library(caret)
library(cowplot)
```

## Intro
This project aims at building predictive models on the [Online News Popularity](https://archive.ics.uci.edu/ml/datasets/Online+News+Popularity#) dataset. Our goal is to build models that predict the number of shares an article receives based on characteristics of the article. This work is important to help writers/companies understand factors that influence article success (as measured by the number of shares an article obtains). This work could also be integrated in the calculation of advertisement space on articles (articles that receive more shares should demand more for ad space). 

The characteristics we will explore include:

+ `weekday`: day of week that article was published (Monday, Tuesday, ...)
+ `num_hrefs`: number of links referenced in article
+ `num_imgs`: number of images in article
+ `num_videos`: number of videos in article
+ `n_tokens_title`: word count of title
+ `n_tokens_content`: word count of article
+ `n_unique_tokens`: number of unique words in article
+ `rate_positive_words` and `rate_negative_words`: rate of positive/negative words among non-neutral tokens
+ `title_subjectivity`: subjectivity of title
+ `avg_positive_polarity` and `avg_negative_polarity`: average polarity of positive/negative words in article

In the model building section, we explore 4 unique models: OLS, logarithmic linear regression, random forest tree, and a boosted tree. These models are trained using cross validation on a training dataset that is 80% of the original dataset. These models are compared using metric scores from results of predictions on test set.

Lastly, all summarizations and models are generated after splitting by data channel. For example, on the lifestyle page, all EDA and model building is using data that is filtered by articles that come from the lifestyle channel. Splitting sections this way causes us to generalize descriptions.

The following packages are required:
```{r, eval=FALSE}
library(tidyverse)
library(corrplot)
library(caret)
library(gbm)
library(leaps)
library(caret)
library(cowplot)
```

## Data manipulaton

We read in the online news popularity data and subset the data by data_channel_is*(one of six groups).

```{r 1_readData}
# read entire dataset
  # subset to get the data channel we want
    # remove unnessary columns
df <- read_csv('data/OnlineNewsPopularity.csv') %>%
        filter((!!sym(paste0('data_channel_is_', params$channel))) == 1) %>%
          select(-contains('data_channel_is'))

dim(df)
```

The object `df` now holds a subset of the data according to the specified data channel. Now, we must create new features to consolidate variables and prepare the dataset for exploratory data analysis:

```{r}
# Create a variable that holds day of week
levels = c('Sunday', 'Monday', 'Tuesday', 'Wednesday','Thursday', 'Friday', 'Saturday')
df <- df %>% mutate(weekday = ifelse(weekday_is_monday==1, 'Monday', 
                                ifelse(weekday_is_tuesday==1, 'Tuesday',
                                ifelse(weekday_is_wednesday==1, 'Wednesday',
                                ifelse(weekday_is_thursday==1, 'Thursday', 
                                ifelse(weekday_is_friday==1, 'Friday',
                                ifelse(weekday_is_saturday==1, 'Saturday', 'Sunday'
                                       ))))))) %>%
                mutate(weekday = factor(weekday, levels = levels))


# function to create the popularity column. 
# popularity rating was created with summary stat info (25%, median, and 75%).
# popularity is going to use for EDA.
popularityCol <- function(dataset){
  dataset <- dataset %>% 
    mutate("Popularity" = if_else(shares > 2800, "Very popular",
                           if_else(shares > 1400, "Somewhat popular", 
                            if_else(shares > 946, "Not too popular", "Not at all popular")) 
                           )
    )            
  return(dataset)
}

# Data set using popularityCol function.
df <- popularityCol(df) %>% as_tibble()

# Overwrite popularity column with factor version
df$Popularity <- as.factor(df$Popularity)

# Use ordered function on a factor to order the levels
df$Popularity <- ordered(df$Popularity, levels = c("Not at all popular", "Not too popular", "Somewhat popular", "Very popular"))
```

We now split the data into train and test sets for predictive modeling.

```{r}
set.seed(123)
# split data into train and test sets
train_rows <- sample(nrow(df), nrow(df)*0.7)
trainData <- df[train_rows,]
testData <- df[-train_rows,] 
```

## Summarizations

### Exploratory Data Anaysis

We will use train set for EDA. Here is the summary statistics of targeted variable (shares).

```{r}
# summary statistics
share_stat <- trainData %>% 
                summarise(Count = n(),
                          Min = min(shares), 
                          Q1 = quantile(shares, 0.25),
                          Median = median(shares),
                          Average = mean(shares),
                          Q3 = quantile(shares, 0.75),
                          Max = max(shares),
                          Std.Dev = sd(shares)
                          )

# Display a table of the summary stats.
knitr::kable(share_stat, digits = 2)
```

#### Shares by day of week

```{r}
# Summary statistics by day of week
trainData %>%
  group_by(weekday) %>%
  summarise(total_shares = sum(shares), avg_shares = round(mean(shares)), max_shares = max(shares)) %>%
  knitr::kable()
```

The above table shows a breakdown of total, average, and maximum number of shares for articles published on a specific weekday for this channel. Some channels tend to have more popular days than others.

```{r}
# histogram for day of week vs shares

trainData %>% ggplot(aes(x=weekday, y=shares)) +
        geom_bar(stat="identity", fill = "darkblue") + 
   theme(axis.text.x = element_text(angle = 45, vjust = .75)) +
    labs(x = "Day of week", 
       title = "Day of Week and Total Number of Shares") 
```

While the summary statistic by day of week table shows the breakdown information, the above bar plot visually shows a count of the total number of shares for all articles published on each day of week. If the total number of shares are generally higher during week days than weekends, we can infer that people are less likely to involve in reading and sharing the articles during the weekends and more likely to relax and do more leisure activities.   

#### Shares by popularity

```{r message=FALSE}
# Summary statistics by popularity
trainData %>% 
  group_by(Popularity) %>%
  summarise(Total_shares = sum(shares), Avg_shares = round(mean(shares)),                     Median_shares = median(shares), IQR = IQR(shares)) %>%
  knitr::kable()
```

The above table show a summary of the newly created `popularity` variable. If the average score is significantly higher than the median score, and the IQR score is also large, then it means that the shares in the specific level (i.e. Very Popular) are more spread out, but the majority are still around the median scores.  

#### Count of news by popularity over different day of week

```{r message=FALSE}
# Bar plot of weekday by popularity 
ggplot(data = trainData, aes(x = weekday)) +
  geom_bar(aes(fill = as.factor(Popularity))) + 
  labs(x = "Day of week", 
       title = "Day of week by popularity") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  scale_fill_discrete(name = "Popularity") 
            
```

The bar plot above shows a breakdown of the number of articles published vs day of week. We can also see a breakdown of the proportion of popularity of articles for each day. This plot is important to understand if a greater number of shares exhibited by a day of the week is due to more articles being published on that day, or if it is due to that day of week having a direct effect on the number of shares.

For example, if all days have the same number of 'very popular' articles, we could hypothesize that day of week does not have an effect on producing 'very popular' articles.

#### Shares by number of links 

```{r message=FALSE}
# Simple scatter plot
g2 <- trainData %>% ggplot(aes(x=num_hrefs, y=shares)) +
        geom_point(size=2, shape=23) +
        ylim(0, 10000) +
        ggtitle("Number of links") +
  geom_smooth(method = lm)
g2
```

In the above scatter, we compare the number of links in an article to its shares. This plot is motivated by the implementation of Google's [PageRank Algorithm](https://en.wikipedia.org/wiki/PageRank). If the points show an upward trend, the articles with more number of links tend to be more shared. 

#### Number of link by day of week 

```{r message=FALSE}
# Box plot of number of link by day of week
ggplot(trainData, aes(x = weekday, 
                                y = num_hrefs, 
                                fill = weekday)
                      ) +
  geom_boxplot() +
  scale_x_discrete("Day of Week") +
  ggtitle("Numbers link comparison by day of week") +
  scale_fill_brewer(palette = "BuPu") +
  scale_fill_discrete(name = "Day of week") 

```

In the above box plot, we look at how the number of link in an article by day of week. When the median line of weekend are lower than weekday, it shares similar pattern of shares that people tend to spend less time reading articles and tend to link less during the weekends. However, if the median line of the weekends are higher, then it infer different story and we may infer that people are more likely to spend time and make effort to link the articles during the weekends since shares are easier to acquire than links.

#### Number of words in the title and content

```{r message=FALSE}
# scatter plot of Number of words in the title
g3 <- ggplot(data = trainData, aes(x =  n_tokens_title, 
                      y = shares)) +
      geom_point(alpha = 0.50) + 
      ggtitle("Word count in the title") +
  geom_smooth(method = lm)


# scatter plot of Number of words in the content
g4 <- ggplot(data = trainData, aes(x =  n_tokens_content, 
                      y = shares)) +
      geom_point(alpha = 0.50) + 
      ggtitle("Word count in the content") +
  geom_smooth(method = lm)

plot_grid(g3, g4,  labels = c('A', 'B'))   
```

Scatter plot A shows the number of words in the title compared to the number of shares. Scatter plot B shows a comparison of the number of words that appear in the article compared to the number of shares. These information will be very useful for the writers because people tend to read articles based on the catch title and length of the articles. If a bell shape appears in the plot A, we can embrace the idea that articles with not too long or short number of words in title are more shared and perhaps a quadratic relationship is appropriate for this variable. Also, in plot B, if the data exhibits a slightly negative slops, we can infer that people tend to share more articles with shorter contents and perhaps a negative linear relationship is appropriate.  

#### Unique words count

```{r message=FALSE}
# scatter plot of Unique words count
ggplot(data = trainData, aes(x =  n_unique_tokens, 
                      y = shares)) +
      geom_point(alpha = 0.50) + 
      ggtitle("Unique Word count") +
      ylim(0, 10000) + 
      geom_smooth(method = lm)
```

Above, we see a plot of the number of unique words in each article compared to the number of shares. Across channels, there does not seem to be a strong linear correlation between these two variables. However, this variable may have significance in tree based models.

#### Number of image and video

```{r message=FALSE}
# scatter plot of Number of words in the content
g5 <- ggplot(data = trainData, aes(x =  num_imgs, 
                      y = shares)) +
      geom_point(alpha = 0.50) +
      ggtitle("Number of image") +
      ylim(0, 10000) +
      geom_smooth(method = lm)

# Scatter plot of Number of words in the content
g6 <- ggplot(data = trainData, aes(x =  num_videos, 
                      y = shares)) +
      geom_point(alpha = 0.50) + 
      ggtitle("Number of video") +
      ylim(0, 10000)+
      geom_smooth(method = lm)

plot_grid(g5, g6,  labels = c('C', 'D')) 
```

In these scatter plots, we compare the number of images/videos to the number of shares. As a reader, when I see interesting images and videos in the article, I have tendency to read and share the articles. However, it may be depending on the topic of the article or numbers of images/videos. If the points display positive trend, then the articles with more images/ videos tend to be shared more. Contrarily, if the points display downward trend, then the articles with more images/videos tend to be shared less. 

#### Number of keywords

```{r message=FALSE}
# Scatter plot of number of keywords
ggplot(data = trainData, aes(x =  num_keywords, 
                      y = shares)) +
      geom_point(alpha = 0.50) + 
      ggtitle("Number of keywords") +
      geom_smooth(method = lm)
```

In above plot, we compare the number of keywords that appear in an article to the number of shares. Across channels, there does not appear to be much linear correlation between the two variables. In other words, the number of keywords may not be the significant factor to the readers, but may be the topic or the name of the keywords are more important aspect since readers often search keywords of their interests to find articles.  

#### Number of positive and negative words rate

```{r message=FALSE}
# Scatter plot of positive words rate
g7 <- ggplot(data = trainData, aes(x =  rate_positive_words, 
                      y = shares)) +
      geom_point(alpha = 0.50) +
      ggtitle("Positive words rate") +
    geom_smooth()


# Scatter plot of and negative words rate
g8 <- ggplot(data = trainData, aes(x =  rate_negative_words, 
                      y = shares)) +
      geom_point(alpha = 0.50) +
      ggtitle("Negative words rate") +
    geom_smooth()

plot_grid(g7, g8,  labels = c('A', 'B')) 
```

In above scatter plots, we compare the positive/negative rate of words that appear in an article to the number of shares of an article. If the positive relationships show in each plot, then articles with more  positive or negative rate of words tend to be shared more frequently. Vice versa, if the negative relationships show in each plot, then articles with  more positive or negative rate of words tend to be shared less frequently. The writers would benefit from the above information whether more positive or negative words should be included across different channels. 

#### Title subjectivity

```{r message=FALSE}
# Scatter plot of title subjectivity
ggplot(data = trainData, aes(x =     title_subjectivity, 
                      y = shares)) +
      geom_point(alpha = 0.50) +
      ggtitle("Title subjectivity") +
      geom_smooth()
```

In the title subjectivity scatter plot, we compare title subjectivity to the number of shares. Across channels, there does not seem to be a strong linear correlation between the two variables. Based on the results, we infer that the subjectivity (such as tone of voice and emotions) in the title may not influence people in terms of sharing articles. 

#### Number of average positive and negative polarity

```{r}
# Scatter plot of average positive polarity
g9 <- ggplot(data = trainData, aes(x =  avg_positive_polarity, 
                      y = shares)) +
      geom_point(alpha = 0.50) +
      ggtitle("Average positive polarity") +
      geom_smooth(method = lm)

# Scatter plot of average negative polarity
g10 <- ggplot(data = trainData, aes(x =  avg_negative_polarity, 
                      y = shares)) +
      geom_point(alpha = 0.50) +
      ggtitle("Average negative polarity") +
      geom_smooth(method = lm)

plot_grid(g9, g10,  labels = c('A', 'B')) 
```

In the above plots, we compare the average polarity of positive/negative words that appear in the article to the number of shares. If there is a positive relationship between the average positive polarity and the number of shares, then articles with more positive polarity tend to be shared more frequently. Yet, if there is a negative relationship between the two variables, then articles with more positive polarity are less likely to be shared. Same interpretation applied to the relationship between average negative polarity and the number of shares that depending on the positive or negative trend, articles with more negative polarity tend to be shared more often or less likely. "Polarity" conveys the idea that words have a very strong opinion. Therefore, higher average positive polarity refers that articles carry extreme positive opinions. 

#### Correlation with numeric variables

```{r}
# Correlation plot
df_tmp <- trainData %>% select(c('n_tokens_title', 
                          'n_tokens_content',
                          'n_unique_tokens',
                          'num_hrefs',
                          'num_imgs',
                          'num_videos',
                          'num_keywords',
                          'is_weekend',
                          'rate_positive_words',
                          'rate_negative_words',
                          'title_subjectivity',
                          'avg_positive_polarity',
                          'avg_negative_polarity',
                          'shares',))
corrplot(cor(df_tmp), type = 'lower', diag = FALSE)

```

Above shows the correlation matrix for the numerical variables we explore in EDA. Shares is the bottom row. We use this plot to find the variables that might have weak correlation with shares and make sure to include these in our model building phase.

## Modeling

### Linear Regression

Linear regression is one of the commonly used supervised methods for modeling and useful tool for predicting a quantitative response on the basis of a single or multiple predictor variables. The idea of linear regression is that the model finds the best fit line between predictor variables and response variable, minimizing sum of squared errors. Also, Linear Regression is flexible in which to include many explanatory variables, higher order terms, and/or interaction, so we can see the effect the combinations on prediction.  Here we will fit two different linear regression models.   

#### Linear model 1

The first linear regression model will have predictors selected by EDA. we will use repeated cross-validation with 10 folder and will find the RMSE and R2.

```{r message=FALSE}
# Linear model 1 with train set 
set.seed(10)
lm.fit1 <- train(shares ~ n_tokens_title + n_tokens_content +
                   num_hrefs + num_imgs + num_videos + num_keywords +
                   rate_positive_words + rate_negative_words + title_subjectivity +
                   n_unique_tokens + avg_positive_polarity + avg_negative_polarity +
                   weekday + I(n_tokens_title^2) + I(n_unique_tokens^2) +
                   I(num_keywords ^2) + I(avg_positive_polarity^2) +
                    I(num_hrefs^2), 
                 data = trainData,
                 method="lm",
                 preProcess = c("center","scale"),
                 trControl = trainControl(method = "repeatedcv", 
                                          number = 10, 
                                          repeats = 3))

lm.fit1
```

First linear model has an RMSE of `r round(lm.fit1$results$RMSE,2)`.  

#### Linear model 2 - Logarithmic Linear Regression

Now, let's look at regresssing on the log-transformed target variable of shares.

```{r}
# Linear model 2 with train set
# Get all numeric columns
train_df <- trainData[ ,unlist(lapply(trainData, is.numeric))]
test_df <- testData[ ,unlist(lapply(testData, is.numeric))]

# code used to get regression variables
## fit using forward selection
#forward <- regsubsets(log(shares) ~ .,
#                      data = train_df,
#                      nvmax = 10,
#                      method = "forward")
## summary
#mod_summary <- summary(forward)

# Train model
lm.fit2 <- train(log(shares) ~ n_tokens_content + num_hrefs + average_token_length + 
                        num_keywords + kw_min_min + kw_max_avg + kw_avg_avg + 
                        is_weekend + LDA_04 + global_subjectivity, 
                        data = train_df,
                        method="lm",
                        preProcess = c("center","scale"),
                        trControl = trainControl(method = "repeatedcv", 
                                                 number = 10, 
                                                 repeats = 3))

lm.fit2
```

Second linear model has an RMSE of `r round(lm.fit2$results$RMSE,2)`.

### Ensemble Tree-based model

Tree-based method splits up response variable into subsets based on their relationship to one or more predictor variables. Because it is easy to understand and interpret output and no statistical assumptions is necessary, regression tree (continuous variable) and classification tree (group membership) are commonly used. However, these models are influenced vastly even there is a small changes in the data. Therefore, Ensemble tree-based model are highly used in machine learning to gain strong prediction even though it loses interpretability because it gets rid of variation of data set to data set and average of predication for a final prediction. In this project, we will use the two of most commonly used ensemble methods: Random Forests and Boosting. 

#### Random Forest Model

Random Forest method,  which shares the idea of bagging, but extends the idea and only include random subset of predictors for each bootstrap sample/tree fit instead of including every predictor in each of the tree.  In doing so, one or two good predictors won’t dominate the tree fit. Random forest method use mtry as tuning parameter and the value of this parameter for regression models is commonly obtained using  *m* = *p*/3 formula (p as the number of predictors). Below, you will see the result of training with the random forest method. By choosing randomly selected subset of predictors in each tree, we will possibly reduce the correlation and gain stronger prediction.

```{r, cache=TRUE}
# Tuning parameter is mtry, use values of 1,..,4
rfFit <- train(shares ~ n_tokens_title + n_tokens_content+
                 n_unique_tokens+avg_positive_polarity+
                 avg_negative_polarity + num_hrefs +  num_imgs +
                 num_videos + num_keywords + title_sentiment_polarity +
                 rate_positive_words + rate_negative_words +
                 title_subjectivity + weekday,
                 data = trainData, 
                 method = "rf", 
                 trControl = trainControl(method = "cv", number = 5),
                 preProcess = c("center", "scale"),
                 tuneGrid = data.frame(mtry = (1:4)))

# Re-train using best hyperparameter value
rfFit <- train(shares ~ n_tokens_title + n_tokens_content+
                 n_unique_tokens+avg_positive_polarity+
                 avg_negative_polarity + num_hrefs +  num_imgs +
                 num_videos + num_keywords + title_sentiment_polarity +
                 rate_positive_words + rate_negative_words +
                 title_subjectivity + weekday,
                 data = trainData, 
                 method = "rf", 
                 trControl = trainControl(method = "cv", number = 5),
                 preProcess = c("center", "scale"),
                 tuneGrid = data.frame(rfFit$bestTune))
rfFit
```

The random forest model has an RMSE of `r round(min(rfFit$results$RMSE),2)`.
  
#### Boosted Tree Model

The boosted tree is another tree-based regression model. This model aims to predict the residuals between the number of shares (target variable) of each observation and the average number of shares. To do this, the model builds smaller trees of specified depth that add/subtract to predictions with the hope that predictions move closer to their residual values. To prevent overfitting of data, trees are constricted to a 'shrinkage' parameter (that takes a value between 0 and 1) which limits the amount of boosting on predictions. 

For a really good video explanation, watch [this](https://www.youtube.com/watch?v=3CC4N4z3GJc).

```{r}
# Get all numeric columns
train_df <- trainData[ ,unlist(lapply(trainData, is.numeric))]
test_df <- testData[ ,unlist(lapply(testData, is.numeric))]

# Declare grid of values to test in cross validation
## Code retrieved from https://topepo.github.io/caret/model-training-and-tuning.html
gbmGrid <-  expand.grid(interaction.depth = c(1, 5, 9), # complexity of tree
                        n.trees = c(25, 50, 100, 150, 200), # number of iterations (i.e. trees)
                        shrinkage = 0.1, # learning rate
                        n.minobsinnode = 20) # minimum number of samples in a node to commence splitting

# Train using crossvalidation, print out best fitting parameters
boostFit <- train(shares ~ .,
                data = train_df,
                method = "gbm",
                trControl = trainControl("cv", number = 5),
                verbose = FALSE,
                tuneGrid = gbmGrid
                )

# Re-train using best hyperparameter value
boostFit <- train(shares ~ .,
                data = train_df,
                method = "gbm",
                trControl = trainControl("cv", number = 5),
                verbose = FALSE,
                tuneGrid = expand.grid(boostFit$bestTune)
                )
boostFit
```

The boosted tree model has an RMSE of `r round(min(boostFit$results$RMSE),2)`.

## Comparison

Now, we compare the 4 models developed above. Each model was evaluated on a test data set and Root MSE values(roof of test prediction error) were compared.

```{r}
# Predict on test data
predLm1 <- predict(lm.fit1, newdata = testData)
predLm2 <- predict(lm.fit2, newdata = test_df)
rfPred <- predict(rfFit, newdata = testData)
boostPred <- predict(boostFit, newdata = test_df)

# Calculate RMSE
rmseLm1 <- sqrt(mean((predLm1 - test_df$shares)^2))
rmseLm2 <- sqrt(mean((predLm2 - test_df$shares)^2))
rfMSE <- sqrt(mean((rfPred - test_df$shares)^2))
boostRMSE <- sqrt(mean((boostPred - test_df$shares)^2))

rmseTotal <- data.frame('Linear Regression Model 1' = rmseLm1, 
                   'Linear Regression Model 2' = rmseLm2, 
                   'Random Forest Model' = rfMSE, 
                   'Boosting Model' = boostRMSE)

knitr::kable(t(rmseTotal),
               digits=2,
               caption="Summary Table of RMSE score",
               col.names = "RMSE")
```

```{r message=FALSE}
# Finding the best model for each channel
winningModel <- ifelse((rmseLm1 < rmseLm2) & (rmseLm1 < rfMSE) & (rmseLm1 < boostRMSE), 'Linear Model 1', 
       ifelse((rmseLm2 < rfMSE) & (rmseLm2 < boostRMSE), 'Linear Model 2',
       ifelse(rfMSE < boostRMSE, 'Random Forest', 'Boosted Tree')))
lowestrmse <- min(rmseLm1, rmseLm2, rfMSE, boostRMSE)
paste('The Winning Model is:', paste0(winningModel, '!'), 'Its RMSE value is', round(lowestrmse, 2))
```

The winning model is `r winningModel`. Its RMSE value is `r round(lowestrmse, 2)`.